\mainmatter

```{r ch1-setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Overview of common visualizations and how to read them {#reading-visualizations}

```{r, include=FALSE}

source('data/inclusiveness_index/load_inclusiveness_index.R')

inclusive.levels <- c("High", "Medium-High", "Medium", "Medium-Low", "Low", "No data")

inclusiveness_index <- inclusiveness_index %>%
    mutate(Index.categories.2020 = Index.categories.2020 %>%
           as_factor() %>%
           fct_relevel(inclusive.levels))

library(RColorBrewer)

inclusive.colors <- rev(c("#BBBBBB", brewer.pal(7, "GnBu")[3:7]))
inclusive.colors <- setNames(inclusive.colors, inclusive.levels)

continent.colors <- setNames(brewer.pal(5, "Set2"), sort(unique(inclusiveness_index$Continent)))

library(hexbin)
library(mapproj)

```

This book will cover how to create a variety of visualizations using R. One of the
first things you should do to improve your skills creating visualizations is to
become familiar with the kinds of visualizations that are possible and the different
features of each. 

Effective visualization design relies on a solid understanding
of how data properties, visualization types, and audience characteristics interact
to help people make sense of a visualization.  In this chapter, we'll look at
a series of common visualization types, and we'll break down how each is meant to
be read. Understanding these basic visualization types will create a solid foundation for
communicating your data science work to a broad audience.

## Visualization components

As we discuss different visualizations, we will also be talking about different
components within the visualizations. In figure \@ref(fig:vis-components) below, 
the major components of the visualization are labeled: the main title, the subtitle,
the x axis title, the y axis title, the panel, the horizontal and vertical gridlines,
and the axis labels and tick marks for both axes. Almost all of the visualizations we cover 
in this book will use these basic components.

```{r vis-components, fig.cap='Visualization components, labeled.'}

tibble(x = c(0:4), y=c(0:4)) %>%
  ggplot(aes(x=x,y=y)) +
  labs(title = "Main title",
       subtitle = "Subtitle",
       x = "x axis title", 
       y = "y axis title") +
  coord_cartesian(xlim=c(0,4), ylim=c(0,4)) +
  annotate("text", x = 0, y = 3.7, label = "panel", size = 5.5, hjust=0) + 
  annotate("text", x = .9, y = 3.1, label = "horizontal gridlines", size = 4, hjust=0) +
  annotate("text", x = 2.45, y = 2.2, label = "vertical gridlines", size = 4, hjust=0, angle=90) + 
  annotate("text", x = .25, y = .75, label = "y axis labels and tick marks", size = 4, hjust=0) +
  annotate("segment", x = .2, xend = -.1, y = .745, yend = 1, 
           arrow = arrow(type = "closed", length = unit(.07, "inches"))) +
  annotate("segment", x = .2, xend = -.1, y = .745, yend = 0, 
           arrow = arrow(type = "closed", length = unit(.07, "inches"))) +
  annotate("text", x = 3.5, y = .75, label = "x axis labels and tick marks", size = 4) +
  annotate("segment", x = 3.5, xend = 3, y = .6, yend = 0, 
           arrow = arrow(type = "closed", length = unit(.07, "inches"))) +
  annotate("segment", x = 3.5, xend = 4, y = .6, yend = 0, 
           arrow = arrow(type = "closed", length = unit(.07, "inches"))) + 
  theme_grey(base_size=13) 

```

In this set of basic visualization components, we see two components labeled as an axis. 
These axes are called the x and y axes, and they always appear in these positions: 
the x axis always goes left to right, and the y axis always goes up and down.

## Bar Chart

The bar chart is possibly the most common type of visualization. In this type
of visualization, the basic shape being used to represent data values is a rectangle. 
In a traditional bar chart, each rectangle (or bar) has exactly the same width, 
and the height of the bar is representative of some data value. To create a simple
bar chart, the data set should have one column that contains textual (or categorical)
data and one column that contains numerical data. A common way to create these two columns
is to start with one categorical data column and count the number of records for each 
category to create the numerical column.

```{r sample-bar, fig.cap='A sample bar chart.'}

ggplot(
  inclusiveness_index %>% 
    mutate(Continent = Continent %>% as_factor() %>% fct_infreq()),
  mapping = aes(x = Continent)) +
  geom_bar() +
  labs(y="Number of Records")

```

In the sample bar chart above, the categorical variable is displayed on the 
x axis and the numerical variable is displayed on the y axis. This results in a 
classic style of bar chart where each bar has the same width and the heights are 
proportionate to a data value. Each bar has a starting value of zero on the y axis. 
Each bar travels upward from zero and stops at the correct data value. 

When reading this visualization, we are comparing the lengths of bars in order to
understand patterns within the numerical data values from our data set. The power 
of the bar chart lies in how precisely we can detect differences in the
end points of the bars. This is something that people naturally do quite well 
when all of the bars start at the same lowest point, or baseline. 

Bar charts are especially effective if the bars that have small differences in 
lengths appear close to each other. In the above chart, this is accomplished by 
arranging the bars so they appear with the highest data values on the left and 
the lowest data values on the right.

```{r sample-bar-neg, fig.cap='A sample bar chart with both positive and negative values.'}

inclusiveness_index %>% 
  drop_na(Inclusiveness.index.2020) %>%
  group_by(Continent) %>%
  summarize(`Mean Inclusiveness Index` = mean(Inclusiveness.index.2020)) %>%
  mutate(Continent = Continent %>% as_factor() %>% fct_reorder(`Mean Inclusiveness Index`, .desc = TRUE)) %>% 
ggplot(mapping = aes(x = Continent, y = `Mean Inclusiveness Index`)) +
  geom_col()

```

Here is another example of a bar chart where the data values include both positive
and negative values. For data values that are negative, the bar travels downward 
from zero and stops at the correct data value. The x axis title and labels appear 
at the bottom of the panel, below the lowest data values.

For stylistic reasons, bar charts may also appear with the bars oriented horizontally
instead of vertically. In that case, each bar will have the same height, and the widths
of the bars will vary based on the data values. The text (or categorical) axis will then
be the y axis, and the numerical axis will be the x axis.

```{r sample-bar-horizontal, fig.cap='A sample bar chart, with the bars oriented horizonally.'}

inclusiveness_index %>% 
  drop_na(Inclusiveness.index.2020) %>%
  group_by(Continent) %>%
  summarize(`Mean Inclusiveness Index` = mean(Inclusiveness.index.2020)) %>%
  mutate(Continent = Continent %>% as_factor() %>% fct_reorder(`Mean Inclusiveness Index`)) %>% 
ggplot(mapping = aes(y = Continent, x = `Mean Inclusiveness Index`)) +
  geom_col()

```

### Variations

Sometimes, we may have data that would work well for a bar chart, but we may want to
explore a variation that would change the basic shape of the chart. There is one main
variation of a bar chart: the lollipop chart. 

#### Lollipop plot

The lollipop plot is a simple variation on the bar chart. In a lollipop plot,
the bars are replace by a long line with a circle at the end, creating something
that looks like a lollipop. Apart from the different shapes used, the lollipop plot 
works just like a bar chart. The circles draw attention to the data value, but the
lines extending to the axis reinforce the length comparisons.

```{r sample-lollipop, fig.cap='A sample lollipop plot.'}

ggplot(
  inclusiveness_index %>% 
    count(Continent) %>%
    mutate(Continent = Continent %>% as_factor() %>% fct_reorder(n, .desc = TRUE)),
  mapping = aes(x = Continent, y=n)) +
  geom_col(width=.03) +
  geom_point(size=4) +
  labs(y="Number of Records")

```

```{r sample-lolli-neg, fig.cap='A sample lollipop plot with both positive and negative values.'}

inclusiveness_index %>% 
  drop_na(Inclusiveness.index.2020) %>%
  group_by(Continent) %>%
  summarize(`Mean Inclusiveness Index` = mean(Inclusiveness.index.2020)) %>%
  mutate(Continent = Continent %>% as_factor() %>% fct_reorder(`Mean Inclusiveness Index`, .desc = TRUE)) %>% 
ggplot(mapping = aes(x = Continent, y = `Mean Inclusiveness Index`)) +
  geom_col(width=.03) +
  geom_point(size=4) 

```

### Add a variable: Bar charts with color

A simple bar chart includes one categorical variable and one numerical variable.
Sometimes, however, it is useful to explore the patterns in relation to a second
categorical variable. Adding another categorical variable to a bar chart usually
means using color to represent the extra variable.

With a stacked bar chart, the additional variable is used to segment the bars into
separate, colored regions. In this example, the bar for each continent is subdivided
into groups of countries based on their values for Inclusive Index.


```{r sample-stacked-bar, fig.cap='A sample stacked bar chart.'}

ggplot(
  inclusiveness_index %>% 
    count(Continent, Index.categories.2020) %>%
    mutate(Continent = Continent %>% as_factor() %>% fct_reorder(n, .fun = sum, .desc = TRUE)),
  mapping = aes(x = Continent, y = n, fill = Index.categories.2020)) +
  geom_col() +
  labs(y="Number of Records", fill="Inclusiveness Index") +
  scale_fill_manual(values=inclusive.colors)

```

With this stacked bar chart, you can still see the total number of records for each Continent,
but what happens when you try to compare the different segments inside the
bar chart? Starting at the bottom, it seems to work out okay. All of the bars
for the "No data" category start at the same baseline (the x axis), and we can
read these segments like a normal bar chart. But what happens with the "Low" segments
right above them? And the "Medium-Low" segments above those? Every time we have
a group of segments that aren't lined up with each other, we have to try to guess
how tall the bar is in comparison to the other bars in the group. The farther
apart the segments are, the harder it is to make that comparison.

Another variation of the stacked bar chart is the "filled" stacked bar chart.
Instead of using the raw counts to determine the lengths of the bars, in the
filled stacked bar chart, the full bars are all stretched to have the same height,
and each segment becomes the percentage of the records in each bar. (Notice how
the y axis changes from "Number" to "Percentage.") This is useful if the percentages 
matter more than the raw counts, but it doesn't fix any of the concerns with 
comparing different segments without a common baseline.

```{r sample-filled-stacked-bar, fig.cap='A sample filled stacked bar chart.'}

ggplot(
  inclusiveness_index %>% 
    count(Continent, Index.categories.2020) %>%
    mutate(Continent = Continent %>% as_factor() %>% fct_reorder(n, .fun = sum, .desc = TRUE)),
  mapping = aes(x = Continent, y = n, fill = Index.categories.2020)) +
  geom_col(position = "fill") +
  labs(y="Percentage of Records", fill="Inclusiveness Index") +
  scale_fill_manual(values=inclusive.colors)

```

An alternative to the stacked bar chart is called the grouped bar chart. In the
grouped bar chart, every segment starts from the x axis. Each continent forms a
group of bars, and each option of the Inclusiveness Index is a separate bar.

```{r sample-grouped-bar, fig.cap='A sample grouped bar chart.'}

ggplot(
  inclusiveness_index,
  mapping = aes(x = Continent, fill = Index.categories.2020)) +
  geom_bar(position = position_dodge2(preserve="single")) +
  labs(y="Number of Records", fill="Inclusiveness Index") +
  scale_fill_manual(values=inclusive.colors)

```

#### Dumbbell plot

Stacked and grouped bar charts show some of the limitations of bar charts for
making complex comparisons. The rectangles in the bar chart take up a large amount
of space. Think back to the lollipop plot, where it's the circles that directly
represent the data value. Converting bars to something like circles opens up the
ability to make more direct comparisons.

For example, let's say we want to compare two continents more directly: Asia and Europe.

```{r sample-grouped-bar-2-cont, fig.cap='A grouped bar chart focusing on two continents.'}

ggplot(
  inclusiveness_index %>% dplyr::filter(Continent %in% c("Asia", "Europe")),
  mapping = aes(x = Continent, fill = Index.categories.2020)) +
  geom_bar(position = position_dodge2(preserve="single")) +
  labs(y="Number of Records", fill="Inclusiveness Index") +
  scale_fill_manual(values=inclusive.colors)

```

This chart groups all of the segments by continent, which makes it easy to compare
different Index categories within a single continent. What if we want to bring more
attention to the difference between continents for each category? We could always
switch which category is the primary division on the x axis and which is 
represented by color.

```{r sample-grouped-bar-2-cont-switched, fig.cap='The same grouped bar chart with a different arrangement of the categorical variables.'}

ggplot(
  inclusiveness_index %>% dplyr::filter(Continent %in% c("Asia", "Europe")),
  mapping = aes(x = Index.categories.2020, fill = Continent)) +
  geom_bar(position = position_dodge2(preserve="single")) +
  labs(y="Number of Records", x="Inclusiveness Index") +
  scale_fill_manual(values=c(continent.colors["Asia"],continent.colors["Europe"]))

```

This improves our ability to compare the continents directly because the bars are
directly next to each other. The amount of space the bars take up is still pretty
large, though. If we combine this chart with something like a lollipop plot, we
get one last variation: a dumbbell plot.

With a dumbbell plot, we use a circle to represent the data values, just like the
lollipop. Instead of having a line that extends all the way to the axis, though,
we use a line to connect the two dots in each category of Inclusiveness Index.


```{r sample-dumbbell, fig.cap='A dumbbell plot comparing two continents.'}

ggplot(
  inclusiveness_index %>% dplyr::filter(Continent %in% c("Asia", "Europe")) %>%
    count(Continent, Index.categories.2020),
  mapping = aes(x = Index.categories.2020, y=n)) +
  geom_line(aes(group=Index.categories.2020), color="#555555", size=1) +
  geom_point(aes(color = Continent), size=6) +
  labs(y="Number of Records", x="Inclusiveness Index") +
  scale_color_manual(values=c(continent.colors["Asia"],continent.colors["Europe"]))

```

With a dumbbell plot, there are three main trends we can explore in the chart. We
can focus on the green circles to see the pattern in the Asia data values. We can
focus on the blue circles to see the pattern in the Europe data values. Finally, 
we can focus on the lengths of the lines connecting the circles to compare the 
continents at each level of the Inclusiveness Index. This chart time is an efficient
way to compare these data values, but remember that it can be difficult to compare
the lengths of shapes when they don't have the same baseline.

## Scatter Plot

The scatter plot is another common visualization type. This type of visualization 
displays one circle for each record in the data set. The position of the circle is
based on the values of two different numerical variables, one of which is associated
with the x axis and the other of which is associated with the y axis.

```{r sample-scatter, fig.cap='A sample scatter plot.'}

inclusiveness_index %>%
  drop_na(Z.score.Religion, Inclusiveness.index.2020) %>%
  ggplot(aes(x = Z.score.Religion, 
             y = Inclusiveness.index.2020)) +
    geom_point() +
  labs(x = "Z-score, Religion",
       y = "Overall Inclusiveness")

```

This visualization is a way to show a relationship between the two numerical variables 
displayed on the axes. A relationship between the variables means that a change in one
variable would predict a specific kind of change in the other variable. For example,
one kind of relationship is a positive correlation, which means that an increase in
one variable is associated with an increase in the other variable. Points with higher
values on the x axis tend to have higher values on the y axis. Similarly, points
with lower values on one axis tend to have lower values on the other axis.

When two numerical variables have a positive correlation, it shows up on the scatter
plot as a diagonal pattern of circles, from the bottom left corner of the chart to 
the upper right corner of the chart. The closer it looks to a straight line, rather 
than a diffuse pattern, the stronger the relationship is.


```{r}

# TODO: series from more diffuse to perfect correlation

```


There are a few types of patterns that might show up when looking at a scatter plot.
Instead of a positive correlation, the variables could have a negative correlation:
high values of one variable are associated with low values of the other variable. This
relationship shows up on the scatter plot as a diagonal patter from the top left
corner to the bottom right corner.

Both positive and negative correlations are linear relationships - they look like
lines on the chart. There are also nonlinear relationships that look like different
kinds of curves. An exponential relationship looks like a curve that starts
mostly horizontal and then curves up dramatically, ending up almost vertical. 
A logarithmic relationship is a curve that starts mostly vertical and bends
over dramatically, ending up almost horizontal. Other curvilinear patterns of dots
might be better represented by other mathematical functions.

```{r sample-relationships, fig.cap='Different relationships between numerical variables.'}

x <- c(1:50)

pos.lin <- tibble(
  type = "positive correlation",
  x = x,
  y = x
)

neg.lin <- tibble(
  type = "negative correlation",
  x = x,
  y = rev(x)
)

exp <- tibble(
  type = "exponential",
  x = x,
  y = x^2
)

log <- tibble(
  type = "logarithmic",
  x = x,
  y = log(x, base = 2)
)


combined <- bind_rows(pos.lin, neg.lin, exp, log) %>%
  mutate(type = type %>% as_factor() %>% 
           fct_relevel(c("positive correlation", "negative correlation",
                         "exponential", "logarithmic")))

ggplot(combined, aes(x, y)) +
  geom_point() +
  facet_wrap(vars(type), scales = "free")


```

What does it mean to see a shape in a scatter plot? A detectable shape in a scatter 
plot is a suggestion that there might be a statistically powerful relationship
between these variables. The chart, however, is not a substitute for a statistical
analysis. Using statistical analyses to explore the relationship between two variables 
is called modeling. 

Sometimes a scatter plot will be combined with a statistical model to explore
the connection between the data points and an ideal relationship. For example, you
may see a scatter plot where there is a correlation between the variables
combined with a linear model (represented by a straight line drawn on top of the points).

```{r sample-scatter-lm, fig.cap='A sample scatter plot with a linear model overlaid over the points.'}

inclusiveness_index %>%
  drop_na(Z.score.Religion, Inclusiveness.index.2020) %>%
  ggplot(aes(x = Z.score.Religion, 
             y = Inclusiveness.index.2020)) +
    geom_point() +
  geom_smooth(method="lm", formula = 'y ~ x', se = FALSE) +
  labs(x = "Z-score, Religion",
       y = "Overall Inclusiveness")

```

When you see a line on a scatter plot like this, it is showing the linear model
that best represents the relationship between the x and y variables. In this case,
the relationship between the variables is not very strong, so the points look more
like a cloud than the straight line of the linear model.

### Variations

The following charts are variations on a scatter plot. They build on a basic
scatter plot by summarizing the data and making it easier to see patterns.

#### Contour or density plot

Sometimes a dataset is too large for a scatter plot to be effective. With a large
number of data points, there can be too much overlap between the circles to see
the dominant patterns. In this instance, it can be helpful to calculate the density 
of data points across the chart and visualize the density instead of (or in addition to)
the points. This is called a contour or density plot.

```{r sample-contour, fig.cap='A sample contour plot with scatter plot points on top.'}

inclusiveness_index_annual %>%
  drop_na(Z.score.Religion, Inclusiveness.index) %>%
  ggplot(aes(x = Z.score.Religion, 
             y = Inclusiveness.index)) +
    geom_density_2d_filled(contour_var = "count") +
    geom_point(color="white") +
  labs(x = "Z-score, Religion",
       y = "Overall Inclusiveness") +
  coord_equal()

```

#### Binned scatter plot

In a contour plot, the density calculation detects regions of high density in a
scatter plot. Another way of summarizing the distribution of points across the plot
is to divide the plot into an even grid and then to count the points inside each
region. This is often called "binning." Common types of binning are rectangular 
(splitting the plot up using a rectangular grid) and hexagonal (splitting the
plot up using a grid of interlocking hexagons).

```{r sample-bin2d, fig.cap='A sample scatter plot binned with a rectangular grid.'}

inclusiveness_index_annual %>%
  drop_na(Z.score.Religion, Inclusiveness.index) %>%
  ggplot(aes(x = Z.score.Religion, 
             y = Inclusiveness.index)) +
    geom_bin2d(binwidth = c(.25,.25)) +
    geom_point(color="white") +
  labs(x = "Z-score, Religion",
       y = "Overall Inclusiveness") +
  coord_equal()

```

```{r sample-hex, fig.cap='A sample scatter plot binned with a hexagonal grid.'}

inclusiveness_index_annual %>%
  drop_na(Z.score.Religion, Inclusiveness.index) %>%
  ggplot(aes(x = Z.score.Religion, 
             y = Inclusiveness.index)) +
    geom_hex(binwidth=.25) +
    geom_point(color="white") +
  labs(x = "Z-score, Religion",
       y = "Overall Inclusiveness") +
  coord_equal()

```

### Add a variable: Scatter plot with color 

So far, our scatter plots have still only been used to visualize the relationship
between two numerical variables. In some datasets, it can be helpful to consider
how an additional variable interacts with the scatter plot pattern. One way to
incorporate an additional variable is to change the color of the points in the
scatter plot according to the third variable. For example, you can associate the
color of the points with a categorical variable to show whether different subsets
of the points cluster in different parts of the graph.


```{r sample-scatter-color, fig.cap='A sample scatter plot with color categories.'}

inclusiveness_index %>%
  drop_na(Z.score.Religion, Inclusiveness.index.2020) %>%
  ggplot(aes(x = Z.score.Religion, 
             y = Inclusiveness.index.2020,
             color=Continent)) +
    geom_point(size=4) +
  labs(x = "Z-score, Religion",
       y = "Overall Inclusiveness") +
  scale_color_manual(values=continent.colors)

```

In the above chart, the color represents the continent; that is, each continent shows 
as a separate color. We're looking for a relationship between the pattern of the colors
and the spatial pattern of the points. In this chart, the points associated with
Europe do overall seem to cluster in the upper-right corner of the graph, meaning that
on the whole the European countries tend to rate highly on both the Z-score value
for religious inclusiveness and the overall inclusiveness score. African countries
also tend to have high scores for religious inclusiveness, but they don't rate as
highly for overall inclusiveness.

You can also use color to visualize a third numerical variable instead of a 
categorical variable.

```{r sample-scatter-color-grad, fig.cap='A sample scatter plot with a color gradient.'}

extreme <- max(max(inclusiveness_index$Z.score.Gender, na.rm=T),
               abs(min(inclusiveness_index$Z.score.Gender, na.rm=T)))

inclusiveness_index %>%
  drop_na(Z.score.Religion, Inclusiveness.index.2020) %>%
  ggplot(aes(x = Z.score.Religion, 
             y = Inclusiveness.index.2020,
             fill=Z.score.Gender)) +
    geom_point(size=4, shape=21, color="grey50") +
  labs(x = "Z-score, Religion",
       y = "Overall Inclusiveness",
       fill = "Z-score, Gender") +
  scale_fill_distiller(type="div", palette=3, limits = c(-extreme, extreme))

```

In this chart, the values in the "Z-score, Gender" variable are associated with
a color gradient. Positive values of this new variable show in the chart as 
increasingly darker shades of purple. Negative values show as increasingly darker
shades of green.

With any chart, adding more variables runs the risk of creating visual confusion
that makes it harder (not easier) to see interesting patterns. For example,
in the chart above, we see that purples mostly occur on the top half of the chart
and greens on the lower half of the chart. Beyond that, though, it's hard to identify
a strong relationship between the strength of the color values and either of the axes.
If colors are not concentrating in a particular region of the plot, adding a third
variable may not be necessary for this chart.

### Add a variable: Bubble chart

A bubble chart is another way of adding a third variable to a scatter plot. Unlike
adding color to the chart, however, a bubble plot works best when you are adding a
third numerical variable. That's because in a bubble plot, the additional variable
is represented by changing the size of the bubble. Representing a categorical variable
by changing the size of the circle isn't as natural as using different colors. It's
hard to focus on all of the bubbles of a particular size to try to identify clusters.

```{r sample-bubble-cat, fig.cap='A sample bubble chart with categories.', warning=FALSE}

inclusiveness_index %>%
  drop_na(Z.score.Religion, Inclusiveness.index.2020) %>%
  ggplot(aes(x = Z.score.Religion, 
             y = Inclusiveness.index.2020,
             size=Continent)) +
    geom_point(shape=21, alpha=.75, color="black", fill="grey50") +
  labs(x = "Z-score, Religion",
       y = "Overall Inclusiveness") 

```

If you use a numerical variable to size the bubbles, the goal of the chart becomes
similar to when color is used to add a third numerical variable: explore whether
the sizes of the bubbles changes in a meaningful way in relation to the axes.


```{r sample-bubble-grad, fig.cap='A sample bubble chart with a size gradient.'}

inclusiveness_index %>%
  drop_na(Z.score.Religion, Inclusiveness.index.2020) %>%
  ggplot(aes(x = Z.score.Religion, 
             y = Inclusiveness.index.2020,
             size=Z.score.Gender)) +
    geom_point(shape=21, alpha=.75, color="black", fill="grey50") +
  labs(x = "Z-score, Religion",
       y = "Overall Inclusiveness",
       size = "Z-score, Gender")

```

There is one property of bubble charts that is different from scatter plots with 
added color. When you have a variable that has both positive and negative values,
it may be a slightly less natural fit for the size of the bubbles. The size of an
object like a circle is naturally a positive value. A circle doesn't itself have
negative size. If the variable uses negative and positive in an abstract sense,
though, the size of a bubble can still help differentiate low and high values.

With bubble charts, color is still available to display another variable if there
is anything else that might interact with the three numerical variables.

```{r sample-bubble-color, fig.cap='A sample bubble chart with color categories.'}

inclusiveness_index %>%
  drop_na(Z.score.Religion, Inclusiveness.index.2020) %>%
  ggplot(aes(x = Z.score.Religion, 
             y = Inclusiveness.index.2020,
             size=Z.score.Gender,
             fill=Continent)) +
    geom_point(shape=21, alpha=.75, color="black") +
  labs(x = "Z-score, Religion",
       y = "Overall Inclusiveness",
       size = "Z-score, Gender") +
  scale_fill_manual(values=continent.colors)

```

Be cautious, again, with adding too many variables to a single chart. If adding
a variable doesn't reveal anything new about the data, it is probably getting in the
way of a pattern that does exist.

## Line Chart

A line chart is typically used to explore the change in a numerical variable over time.
The variable is measured over a series of time points (for example, years, days, seconds).
The time variable is associated with the x axis of the chart, and the numerical variable
is associated with the y axis. Instead of placing a circle for each record in the
data set, like you would see in a scatter plot, the chart uses a line that travels
through all of the correct data points. Connecting related points with a line
gives the data a sense of continuity and motion. Line charts are often used to look 
for dramatic increases or decreases in the numerical value at particular points in time.


```{r}


inclusiveness_index_annual %>%
  dplyr::filter(Country == "Rwanda") %>%
ggplot(aes(x = Year, 
           y = Inclusiveness.index)) +
  geom_line() +
  labs(title = "Rwanda")

```

Line charts can show change over time for one entity or for many entities. In the example
above, there is a single line representing a country. You can also use a line chart
to show multiple lines at once. For example, if each line is a country, you can display
many countries on the same chart to allow for comparisons between countries as well as
over time.


```{r}

inclusiveness_index_annual %>%
  drop_na(Inclusiveness.index) %>%
ggplot(aes(x = Year, 
           y = Inclusiveness.index,
           group = Country)) +
  geom_line() +
  labs(title = "All Countries")

```

As with the other visualizations we've looked at, it is easy to create a line chart
with so much data that it is hard to see patterns. The goal of the line chart is to
show trends over time, but if all of the lines look flat, or if there are so many
lines that you cannot detect the variations that are there, it may be useful to include
only a small number of the lines at once, or to combine groups of the lines together. 


```{r}

inclusiveness_index_annual %>%
  group_by(Continent, Year) %>%
  summarise(mean.incl.ind = mean(Inclusiveness.index, na.rm=T), .groups="drop") %>%
ggplot(aes(x = Year, 
           y = mean.incl.ind,
           group = Continent)) +
  geom_line() +
  geom_text(data = . %>% dplyr::filter(Year == 2020), 
            aes(label = Continent),
            nudge_y = .03,
            hjust = 0) +
  labs(title = "Average by Continent") +
  scale_x_continuous(limits = c(2016, 2020.3))

```

In this version of the chart, the countries from each continent have been combined
into a single line using the mean of the countries' values for each year. Another way to
reduce the number of lines in the line chart would have been to limit the chart to
the countries from a single continent.

*TODO: this has become about creating visualizations, instead of reading them!!*

Also notice that when you limit the number of lines you use in a line chart, it 
becomes much easier to identify each line with a label. Not every visualization 
includes labels for every data point. For example, in a scatter plot, you may not
label the points if the important information is the overall trend. You may have
the same situation in a line chart if all of the lines are moving in a similar 
direction, or they all experience a dramatic change at the same point. If it is
important to identify specific lines in your chart, however, it is best to use only
a few lines so that it is possible to identify each line directly and trace it
individually across the entire chart.

#### Using circles to highlight the position of the data points

Some line charts include circles along the lines. These circles indicate that the
data set included a value at that point in the line. This can be especially useful
if lines travel through data at irregular time intervals. For example, if the data
is collected every year except that there was a gap of three years in the middle, 
the circles would clearly indicate which years are included in the dataset and which
are not. Normally the lines on a line chart will connect straight from one point in
time to the next point in time, and unless the line turns up or down, it can be
hard to know exactly where the real data values are.

```{r}

inclusiveness_index_annual %>%
  group_by(Continent, Year) %>%
  summarise(mean.incl.ind = mean(Inclusiveness.index, na.rm=T), .groups="drop") %>%
ggplot(aes(x = Year, 
           y = mean.incl.ind,
           group = Continent)) +
  geom_line() +
  geom_point() + 
  geom_text(data = . %>% dplyr::filter(Year == 2020), 
            aes(label = Continent),
            nudge_y = .03,
            hjust = 0) +
  labs(title = "Average by Continent") +
  scale_x_continuous(limits = c(2016, 2020.3))

```

#### Dual encoding names with color

Sometimes line charts use color to differentiate the lines from each other. Each line
gets its own color, and the color legend clarifies the name of the line. Even for
charts where it is important to individually identify the lines, it can quickly
become overwhelming to see a separate color for each line. Even a small number of 
lines can require a set of colors that include colors that are difficult to tell
apart from each other. In additional, it can be confusing and time consuming to have
to consult a legend to identify an individual line. While color may be a nice 
complement to a simple line chart, it is easier to read a line chart when the lines
are labeled directly instead of relying on a legend.

```{r}

inclusiveness_index_annual %>%
  group_by(Continent, Year) %>%
  summarise(mean.incl.ind = mean(Inclusiveness.index, na.rm=T), .groups="drop") %>%
ggplot(aes(x = Year, 
           y = mean.incl.ind,
           group = Continent)) +
  geom_line() +
  geom_text(data = . %>% dplyr::filter(Year == 2020), 
            aes(label = Continent),
            nudge_y = .03,
            hjust = 0) +
  labs(title = "Average by Continent") +
  scale_x_continuous(limits = c(2016, 2020.3))

inclusiveness_index_annual %>%
  group_by(Continent, Year) %>%
  summarise(mean.incl.ind = mean(Inclusiveness.index, na.rm=T), .groups="drop") %>%
ggplot(aes(x = Year, 
           y = mean.incl.ind,
           color = Continent)) +
  geom_line() +
  labs(title = "Average by Continent")


inclusiveness_index_annual %>%
  group_by(Continent, Year) %>%
  summarise(mean.incl.ind = mean(Inclusiveness.index, na.rm=T), .groups="drop") %>%
ggplot(aes(x = Year, 
           y = mean.incl.ind,
           color = Continent)) +
  geom_line(show.legend = FALSE) +
  geom_text(data = . %>% dplyr::filter(Year == 2020), 
            aes(label = Continent),
            nudge_y = .03,
            hjust = 0,
            show.legend = FALSE) +
  labs(title = "Average by Continent") +
  scale_x_continuous(limits = c(2016, 2020.4))

```

### Variations

There is one primary variation of the line chart: the area chart. 

#### Area chart

An area chart looks like a line chart where everything under the line, down to the
x axis, has been colored in. If you took a typical line chart and color everything
in under each line, though, you would have overlapping colors for much of the chart.
It is more common to see a stacked area chart, which follows the same principles as
a stacked bar chart. One line is chosen as the bottom line, and the values for
each additional line are stacked on top of that line. The top most line shows the
overall total of the data in the chart. For this reason, area charts should be used
to show data points that can be added together, like counts.

```{r}
inclusiveness_index_annual %>%
  dplyr::filter(!is.na(Inclusiveness.index)) %>%
  count(Continent, Year) %>%
  mutate(Continent = Continent %>% as_factor() %>% fct_reorder(n, .fun = sum)) %>%
ggplot(aes(x = Year, 
           y = n,
           fill = Continent)) +
  geom_area() +
  scale_fill_manual(values=continent.colors, 
                    breaks =  names(sort(table(inclusiveness_index_annual$Continent[!is.na(inclusiveness_index_annual$Inclusiveness.index)]))))

```

Stacked area charts raise concerns that are similar to the ones raised by stacked
bar charts. The values for the lowest line and the overall total are easy to 
read, but the inner areas do not have a consistent baseline. It can be very difficult
to see if an area has its own decrease at a particular year or if it is just
reacting to a decrease in one of the lines below it.

### Add a variable: Line chart with color

* alternately, you can use color to layer an additional variable on top of the lines
* usually a way to create groups of lines that are all part of the same category
* or that are showing similar data patterns


```{r}

inclusiveness_index_annual %>%
  drop_na(Inclusiveness.index) %>%
ggplot(aes(x = Year, 
           y = Inclusiveness.index,
           group = Country,
           color = Continent)) +
  geom_line(alpha = .75) +
  #geom_line(
  #  data = inclusiveness_index_annual %>% 
  #    group_by(Continent, Year) %>%
  #    summarise(med_incl = median(Inclusiveness.index, na.rm = TRUE), .groups = "drop"),
  #    aes(x = Year, y = med_incl, group = Continent), size = 3, color = "#555555"
  #) +
  #geom_line(
  #  data = inclusiveness_index_annual %>% 
  #    group_by(Continent, Year) %>%
  #    summarise(med_incl = median(Inclusiveness.index, na.rm = TRUE), .groups = "drop"),
  #    aes(x = Year, y = med_incl, color = Continent, group = NULL), size = 2
  #) +
  scale_color_manual(values=continent.colors)

```

```{r}

inclusiveness_index_annual %>%
  drop_na(Inclusiveness.index) %>%
ggplot(aes(x = Year, 
           y = Inclusiveness.index,
           group = Country)) +
  geom_line(alpha = .75, color = "#bbbbbb") +
  geom_line(
    data = . %>%
      dplyr::filter(Continent == "Americas"),
      aes(color = Continent),
    size = 1
  ) +
  scale_color_manual(values=continent.colors, breaks = c("Americas")) +
  theme_bw()

```

## Pie Chart

```{r}

inclusiveness_index %>%
  ggplot(aes(y = "", fill = Index.categories.2020)) +
    geom_bar(position=position_fill()) +
    coord_polar(direction = -1) +
    scale_fill_manual(values = inclusive.colors) +
    theme_void()

```

### Variations

#### Donut chart

```{r}

inclusiveness_index %>%
  count(Continent, Index.categories.2020) %>%
  ggplot(aes(x = 4, fill = Index.categories.2020, y=n)) +
    geom_col(position="fill") +
    coord_polar(theta="y", direction = -1) +
    xlim(c(2.5, 4.5)) +
    scale_fill_manual(values = inclusive.colors) +
    theme_void()


```

#### Waffle chart

### Add a variable: Small multiples

```{r}

inclusiveness_index %>%
  ggplot(aes(y = "", fill = Index.categories.2020)) +
    geom_bar(position=position_fill()) +
    coord_polar(direction = -1) +
    scale_fill_manual(values = inclusive.colors) +
    theme_void() +
    facet_wrap(vars(Continent))

```

## Heat Map

```{r, warning=FALSE}

inclusiveness_index %>%
  pivot_longer(cols = c(Inclusiveness.index.2020, starts_with("Z")), 
               names_to = "index_name",
               values_to = "index_values",
               values_drop_na = TRUE) %>%
  group_by(Index.categories.2020, index_name) %>%
  summarise(mean_incl_ind = mean(index_values), .groups="drop") %>%
  mutate(index_name = index_name %>%
           as_factor() %>%
           fct_relevel(c("Z.score.General.Population", "Z.score.Race", "Z.score.Gender", 
                         "Z.score.Sexual.Orientation", "Z.score.Religion", 
                         "Z.score.Disability", "Inclusiveness.index.2020")) %>%
           fct_rev()) %>%
  #View() %>%
  ggplot(aes(y = index_name, x = Index.categories.2020, fill = mean_incl_ind)) +
    geom_tile() +
    scale_fill_distiller(type="div", palette = 3, 
                         limits=c(-1.2,1.2), 
                         direction = -1) +
  theme_bw()

```

### Variations

#### Circle matrix

```{r}

inclusiveness_index %>%
  pivot_longer(cols = c(Inclusiveness.index.2020, starts_with("Z")), 
               names_to = "index_name",
               values_to = "index_values",
               values_drop_na = TRUE) %>%
  group_by(Index.categories.2020, index_name) %>%
  summarise(mean_incl_ind = mean(index_values), .groups="drop") %>%
  mutate(index_name = index_name %>%
           as_factor() %>%
           fct_relevel(c("Z.score.General.Population", "Z.score.Race", "Z.score.Gender", 
                         "Z.score.Sexual.Orientation", "Z.score.Religion", 
                         "Z.score.Disability", "Inclusiveness.index.2020")) %>%
           fct_rev()) %>%
  #View() %>%
  ggplot(aes(y = index_name, x = Index.categories.2020, fill = mean_incl_ind)) +
    geom_point(shape=21, size=7) +
    scale_fill_distiller(type="div", palette = 3, 
                         limits=c(-1.2,1.2), 
                         direction = -1) +
  theme_bw()

```

### Add a variable: Bubble matrix

```{r}

inclusiveness_index %>%
  pivot_longer(cols = c(Inclusiveness.index.2020, starts_with("Z")), 
               names_to = "index_name",
               values_to = "index_values",
               values_drop_na = TRUE) %>%
  group_by(Index.categories.2020, index_name) %>%
  summarise(mean_incl_ind = mean(index_values), n=n(), .groups="drop") %>%
  mutate(index_name = index_name %>%
           as_factor() %>%
           fct_relevel(c("Z.score.General.Population", "Z.score.Race", "Z.score.Gender", 
                         "Z.score.Sexual.Orientation", "Z.score.Religion", 
                         "Z.score.Disability", "Inclusiveness.index.2020")) %>%
           fct_rev()) %>%
  #View() %>%
  ggplot(aes(y = index_name, x = Index.categories.2020, fill = mean_incl_ind, size=n)) +
    geom_point(shape=21) +
    scale_fill_distiller(type="div", palette = 3, 
                         limits=c(-1.2,1.2), 
                         direction = -1) +
    scale_size_continuous(range = c(3,11)) + 
  theme_bw()

```

## Histogram

Largely one-variable, one-axis plots

```{r}

inclusiveness_index %>%
  drop_na(Inclusiveness.index.2020) %>%
  ggplot(aes(x = Inclusiveness.index.2020)) +
    geom_histogram(binwidth=.09)

```


### Variations

#### Dot plot

```{r}

inclusiveness_index %>%
  drop_na(Inclusiveness.index.2020) %>%
  ggplot(aes(x = Inclusiveness.index.2020)) +
    geom_dotplot(binwidth=.09)

```

#### Frequency Polygon

```{r}

inclusiveness_index %>%
  drop_na(Inclusiveness.index.2020) %>%
  ggplot(aes(x = Inclusiveness.index.2020)) +
    geom_freqpoly(binwidth=.09)

```

#### Density

```{r}

inclusiveness_index %>%
  drop_na(Inclusiveness.index.2020) %>%
  ggplot(aes(x = Inclusiveness.index.2020)) +
    geom_density()

```

#### Box Plot

```{r}

inclusiveness_index %>%
  drop_na(Inclusiveness.index.2020) %>%
  ggplot(aes(y = Inclusiveness.index.2020)) +
    geom_boxplot() +
    scale_x_continuous(limits = c(-2,2))

```

#### Centered dot plot / Bee swarm


```{r}

inclusiveness_index %>%
  drop_na(Inclusiveness.index.2020) %>%
  ggplot(aes(x=0, y = Inclusiveness.index.2020)) +
    geom_dotplot(binaxis = "y", stackdir = "center", binwidth=.09) +
    scale_x_continuous(limits = c(-2,2))

```

#### Violin plot

```{r}

inclusiveness_index %>%
  drop_na(Inclusiveness.index.2020) %>%
  ggplot(aes(x=0, y = Inclusiveness.index.2020)) +
    geom_violin() +
    scale_x_continuous(limits = c(-2,2))

```

### Add a variable: Separate groups on x axis

```{r}

inclusiveness_index %>%
  drop_na(Inclusiveness.index.2020) %>%
  ggplot(aes(x = Continent, y = Inclusiveness.index.2020)) +
    geom_boxplot()

```

```{r}

inclusiveness_index %>%
  drop_na(Inclusiveness.index.2020) %>%
  ggplot(aes(x = Continent, y = Inclusiveness.index.2020)) +
    geom_dotplot(binaxis = "y", stackdir = "center", binwidth=.09)

```

```{r}

inclusiveness_index %>%
  drop_na(Inclusiveness.index.2020) %>%
  ggplot(aes(x = Continent, y = Inclusiveness.index.2020)) +
    geom_violin()

```

### Add a variable: Separate groups with color

```{r}

inclusiveness_index %>%
  drop_na(Inclusiveness.index.2020) %>%
  ggplot(aes(x = Inclusiveness.index.2020, fill = Continent)) +
    geom_histogram(binwidth=.09) +
    scale_color_manual(values = continent.colors)

```

```{r}

inclusiveness_index %>%
  drop_na(Inclusiveness.index.2020) %>%
  ggplot(aes(x = Inclusiveness.index.2020, fill = Continent)) +
    geom_histogram(binwidth=.09, position="dodge") +
    scale_color_manual(values = continent.colors)

```

```{r}

inclusiveness_index %>%
  drop_na(Inclusiveness.index.2020) %>%
  ggplot(aes(x = Inclusiveness.index.2020, color = Continent)) +
    geom_freqpoly(binwidth=.09) +
    scale_color_manual(values = continent.colors)

```



```{r}

inclusiveness_index %>%
  drop_na(Inclusiveness.index.2020) %>%
  ggplot(aes(x = Inclusiveness.index.2020, fill = Continent)) +
    geom_density(alpha = .25) +
  scale_fill_manual(values = continent.colors)

```


## Maps

### Choropleth 

```{r}

map_incl_ind <- inclusiveness_index %>%
  mutate(Country = case_when(
    Country == "Antigua and Barbuda" ~ "Antigua",
    Country == "Bonaire, Saint Eustatius and Saba" ~ "Bonaire",
    Country == "Cote d'Ivoire" ~ "Ivory Coast",
    Country == "East Timor" ~ "Timor-Leste",
    Country == "Palestina" ~ "Palestine",
    Country == "Saint Kitts and Nevis" ~ "Saint Kitts",
    Country == "Saint Vincent and the Grenadines" ~ "Saint Vincent",
    Country == "Saint-Barthalemy" ~ "Saint Barthelemy",
    Country == "Saint-Martin" ~ "Saint Martin",
    Country == "Trinidad and Tobago" ~ "Trinidad",
    Country == "United Kingdom" ~ "UK",
    Country == "United States" ~ "USA",
    Country == "Vatican City" ~ "Vatican",
    TRUE ~ Country
  )) %>%
  bind_rows(
    inclusiveness_index %>% dplyr::filter(
      Country %in% c("Antigua and Barbuda", "Bonaire, Saint Eustatius and Saba",
                     "Saint Kitts and Nevis", "Saint Vincent and the Grenadines",
                     "Trinidad and Tobago")
    ) %>% mutate(Country = case_when(
      Country == "Antigua and Barbuda" ~ "Barbuda",
      Country == "Bonaire, Saint Eustatius and Saba" ~ "Sint Eustatius",
      Country == "Saint Kitts and Nevis" ~ "Nevis",
      Country == "Saint Vincent and the Grenadines" ~ "Grenadines",
      Country == "Trinidad and Tobago" ~ "Tobago",
      TRUE ~ Country
      )
    ),
    inclusiveness_index %>% dplyr::filter(
      Country %in% c("Bonaire, Saint Eustatius and Saba")
    ) %>% mutate(Country = case_when(
      Country == "Bonaire, Saint Eustatius and Saba" ~ "Saba",
      TRUE ~ Country
      )
    )
  ) 

map_incl_ind %>%
  left_join(map_data("world2"), by = c("Country" = "region")) %>%
  ggplot() + 
    geom_polygon(aes(x = long,
                     y = lat,
                     group=group,
                     fill=Inclusiveness.index.2020)) +
    coord_map() +
    scale_fill_distiller(type="div", palette = 3, 
                         limits=c(-1.2,1.2), 
                         direction = 1)


```


### Proportional Symbol Map

```{r, warning=FALSE}

map_incl_ind %>%
  drop_na(Inclusiveness.index.2020) %>%
  left_join(
    map_data("world2") %>% 
      group_by(region) %>%
      summarise(x = mean(long),
                y = mean(lat)), 
    by = c("Country" = "region")) %>%
  ggplot() +
    geom_polygon(
      data = map_data("world2"),
      aes(x = long,
          y = lat,
          group=group),
      fill = "grey85",
      color = "grey70") +
    coord_map() +
    geom_point(aes(x = x, y = y, 
                   fill = Index.categories.2020, 
                   size = Inclusiveness.index.2020),
               shape = 21,
               alpha = .75) +
    scale_fill_manual(values = inclusive.colors[names(inclusive.colors) != "No data"]) +
    scale_y_continuous(limits = c(-360,360))

```
