\mainmatter

```{r ch1-setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Overview of common visualizations and how to read them {#reading-visualizations}

```{r, include=FALSE}

source('data/inclusiveness_index/load_inclusiveness_index.R')

inclusiveness_index <- inclusiveness_index %>%
    mutate(Index.categories.2020 = Index.categories.2020 %>%
           as_factor() %>%
           fct_relevel(c("High", "Medium-High", "Medium", "Medium-Low", "Low", "No data")))

```

```{r, eval=FALSE}

inclusiveness_index_annual %>%
  drop_na(Inclusiveness.index) %>%
  ggplot(aes(x = Year, y = Inclusiveness.index, color = Continent)) +
    geom_jitter()

```

This book will cover how to create a variety of visualizations using R. One of the
first things you should do to improve your skills creating visualizations is to
become familiar with the kinds of visualizations that are possible and the different
features of each. 

Effective visualization design relies on a solid understanding
of how data properties, visualization types, and audience characteristics interact
to help people make sense of a visualization.  In this chapter, we'll look at
a series of common visualization types, and we'll break down how each is meant to
be read. Understanding these basic visualization types will create a solid foundation for
communicating your data science work to a broad audience.

## Visualization components

As we discuss different visualizations, we will also be talking about different
components within the visualizations. In figure \@ref(fig:vis-components) below, 
the major components of the visualization are labeled: the main title, the subtitle,
the x axis title, the y axis title, the panel, the horizontal and vertical gridlines,
and the axis labels and tick marks for both axes. Almost all of the visualizations we cover 
in this book will use these basic components.

```{r vis-components, fig.cap='Visualization components, labeled.'}

tibble(x = c(0:4), y=c(0:4)) %>%
  ggplot(aes(x=x,y=y)) +
  labs(title = "Main title",
       subtitle = "Subtitle",
       x = "x axis title", 
       y = "y axis title") +
  coord_cartesian(xlim=c(0,4), ylim=c(0,4)) +
  annotate("text", x = 0, y = 3.7, label = "panel", size = 5.5, hjust=0) + 
  annotate("text", x = .9, y = 3.1, label = "horizontal gridlines", size = 4, hjust=0) +
  annotate("text", x = 2.45, y = 2.2, label = "vertical gridlines", size = 4, hjust=0, angle=90) + 
  annotate("text", x = .25, y = .75, label = "y axis labels and tick marks", size = 4, hjust=0) +
  annotate("segment", x = .2, xend = -.1, y = .745, yend = 1, 
           arrow = arrow(type = "closed", length = unit(.07, "inches"))) +
  annotate("segment", x = .2, xend = -.1, y = .745, yend = 0, 
           arrow = arrow(type = "closed", length = unit(.07, "inches"))) +
  annotate("text", x = 3.5, y = .75, label = "x axis labels and tick marks", size = 4) +
  annotate("segment", x = 3.5, xend = 3, y = .6, yend = 0, 
           arrow = arrow(type = "closed", length = unit(.07, "inches"))) +
  annotate("segment", x = 3.5, xend = 4, y = .6, yend = 0, 
           arrow = arrow(type = "closed", length = unit(.07, "inches"))) + 
  theme_grey(base_size=13) 

```

In this set of basic visualization components, we see two components labeled as an axis. 
These axes are called the x and y axes, and they always appear in these positions: 
the x axis always goes left to right, and the y axis always goes up and down.

## Bar Chart

The bar chart is possibly the most common type of visualization. In this type
of visualization, the basic shape being used to represent data values is a rectangle. 
In a traditional bar chart, each rectangle (or bar) has exactly the same width, 
and the height of the bar is representative of some data value. To create a simple
bar chart, the data set should have one column that contains textual data and one
column that contains numerical data. (Another common way to create a bar chart
is to start with one textual data column and count each instance of each text
value to create the numerical column.)

```{r sample-bar, fig.cap='A sample bar chart.'}

inclusiveness_index_annual %>% 
  drop_na(Inclusiveness.index) %>%
  group_by(Continent) %>%
  summarize(`Mean Inclusiveness Index` = mean(Inclusiveness.index)) %>%
  mutate(Continent = Continent %>% as_factor() %>% fct_reorder(`Mean Inclusiveness Index`, .desc = TRUE)) %>% 
ggplot(mapping = aes(x = Continent, y = `Mean Inclusiveness Index`)) +
  geom_col()

```

In the sample bar chart above, we see a classic style of bar chart where each
bar has the same width and the heights are proportionate to a data value. Each 
bar has a starting value of zero on the y axis. For data values that are positive,
the bar travels upward from zero and stops at the correct data value. For data values
that are negative, the bar travels downward from zero and stops at the correct
data value. The x axis title and labels appear at the bottom of the panel, below
the lowest data values.

When reading this visualization, we are comparing the lengths of bars in order to
understand patterns within the numerical data values from our data set. The power 
of the bar chart lies in how precisely we can detect differences in the
lengths of bars. This is something that people naturally do quite well. Bar charts 
are especially effective if the bars that have small differences in lengths appear 
close to each other. In the above chart, this is accomplished by arranging the bars 
so they appear with the highest data values on the left and the lowest data values 
on the right.

For stylistic reasons, bar charts may also appear with the bars oriented horizontally
instead of vertically. In that case, each bar will have the same height, and the widths
of the bars will vary based on the data values. The text (or categorical) axis will then
be the y axis, and the numerical axis will be the x axis.

```{r sample-bar-horizontal, fig.cap='A sample bar chart, with the bars oriented horizonally.'}

inclusiveness_index_annual %>% 
  drop_na(Inclusiveness.index) %>%
  group_by(Continent) %>%
  summarize(`Mean Inclusiveness Index` = mean(Inclusiveness.index)) %>%
  mutate(Continent = Continent %>% as_factor() %>% fct_reorder(`Mean Inclusiveness Index`)) %>% 
ggplot(mapping = aes(y = Continent, x = `Mean Inclusiveness Index`)) +
  geom_col()

```

### Variations

#### Stacked and grouped bar charts

#### Dot plots

#### Dumbbell plots


## Scatter Plot

scatter plot / scatter plot with color / bubble chart / countour plot

```{r}

inclusiveness_index %>%
  drop_na(Z.score.Religion, Inclusiveness.index.2020) %>%
  ggplot(aes(x = Z.score.Religion, 
             y = Inclusiveness.index.2020)) +
    geom_point()

```


## Line Chart

line chart / area chart

```{r}

inclusiveness_index_annual %>%
  group_by(Year, Continent) %>%
  summarise(mean_incl_ind = mean(Inclusiveness.index, na.rm = TRUE)) %>%
  dplyr::filter(Continent == "Europe") %>%
ggplot(aes(x = Year, 
           y = mean_incl_ind)) +
  geom_line()

inclusiveness_index_annual %>%
  group_by(Year, Continent) %>%
  summarise(mean_incl_ind = mean(Inclusiveness.index, na.rm = TRUE)) %>%
  ggplot(aes(x = Year, 
             y = mean_incl_ind,
             color = Continent)) +
    geom_line()

```

## Pie Chart

pie chart / donut chart

```{r}

# library(RColorBrewer)
# brewer.pal(5, "Blues")

inclusiveness_index %>%
  #dplyr::filter(Continent == "Americas") %>%
  ggplot(aes(y = "", fill = Index.categories.2020)) +
    geom_bar(position=position_fill()) +
    coord_polar(direction = -1) +
    #scale_fill_brewer(type="seq", palette = 1, direction = -1) +
    scale_fill_manual(values = c("#08519C","#3182BD","#6BAED6","#9ECAE1","#C6DBEF","#EEEEEE")) +
    theme_void() +
    facet_wrap(vars(Continent))

```


## Heat Map

heat map / matrix / circles with color and size

```{r}

inclusiveness_index %>%
  pivot_longer(cols = c(Inclusiveness.index.2020, starts_with("Z")), 
               names_to = "index_name",
               values_to = "index_values",
               values_drop_na = TRUE) %>%
  group_by(Index.categories.2020, index_name) %>%
  summarise(mean_incl_ind = mean(index_values)) %>%
  mutate(index_name = index_name %>%
           as_factor() %>%
           fct_relevel(c("Z.score.General.Population", "Z.score.Race", "Z.score.Gender", 
                         "Z.score.Sexual.Orientation", "Z.score.Religion", 
                         "Z.score.Disability", "Inclusiveness.index.2020")) %>%
           fct_rev()) %>%
  #View() %>%
  ggplot(aes(y = index_name, x = Index.categories.2020, fill = mean_incl_ind)) +
    geom_tile() +
    scale_fill_distiller(type="div", palette = 3, 
                         limits=c(-1.2,1.2), 
                         direction = -1) +
  theme_bw()

```


## Histogram

histogram / density

```{r}

inclusiveness_index %>%
  drop_na(Inclusiveness.index.2020) %>%
  ggplot(aes(x = Inclusiveness.index.2020)) +
    geom_histogram()

```


## Box Plot

box plot / violin plot / bee swarm

```{r}

ggplot(inclusiveness_index, 
       aes(group = Z.score.Disability, 
           y = Inclusiveness.index.2020)) +
  geom_boxplot()

```


## Maps

choropleth / proportional symbol map

```{r}

map_incl_ind <- inclusiveness_index %>%
  mutate(Country = case_when(
    Country == "Antigua and Barbuda" ~ "Antigua",
    Country == "Bonaire, Saint Eustatius and Saba" ~ "Bonaire",
    Country == "Cote d'Ivoire" ~ "Ivory Coast",
    Country == "East Timor" ~ "Timor-Leste",
    Country == "Palestina" ~ "Palestine",
    Country == "Saint Kitts and Nevis" ~ "Saint Kitts",
    Country == "Saint Vincent and the Grenadines" ~ "Saint Vincent",
    Country == "Saint-Barthalemy" ~ "Saint Barthelemy",
    Country == "Saint-Martin" ~ "Saint Martin",
    Country == "Trinidad and Tobago" ~ "Trinidad",
    Country == "United Kingdom" ~ "UK",
    Country == "United States" ~ "USA",
    Country == "Vatican City" ~ "Vatican",
    TRUE ~ Country
  )) %>%
  bind_rows(
    inclusiveness_index %>% dplyr::filter(
      Country %in% c("Antigua and Barbuda", "Bonaire, Saint Eustatius and Saba",
                     "Saint Kitts and Nevis", "Saint Vincent and the Grenadines",
                     "Trinidad and Tobago")
    ) %>% mutate(Country = case_when(
      Country == "Antigua and Barbuda" ~ "Barbuda",
      Country == "Bonaire, Saint Eustatius and Saba" ~ "Sint Eustatius",
      Country == "Saint Kitts and Nevis" ~ "Nevis",
      Country == "Saint Vincent and the Grenadines" ~ "Grenadines",
      Country == "Trinidad and Tobago" ~ "Tobago",
      TRUE ~ Country
      )
    ),
    inclusiveness_index %>% dplyr::filter(
      Country %in% c("Bonaire, Saint Eustatius and Saba")
    ) %>% mutate(Country = case_when(
      Country == "Bonaire, Saint Eustatius and Saba" ~ "Saba",
      TRUE ~ Country
      )
    )
  ) 

map_incl_ind %>%
  left_join(map_data("world2"), by = c("Country" = "region")) %>%
  ggplot() + 
    geom_polygon(aes(x = long,
                     y = lat,
                     group=group,
                     fill=Inclusiveness.index.2020)) +
    coord_map() +
    scale_fill_distiller(type="div", palette = 3, 
                         limits=c(-1.2,1.2), 
                         direction = -1)

map_incl_ind %>%
  drop_na(Inclusiveness.index.2020) %>%
  left_join(
    map_data("world2") %>% 
      group_by(region) %>%
      summarise(x = mean(long),
                y = mean(lat)), 
    by = c("Country" = "region")) %>%
  ggplot() +
    geom_polygon(
      data = map_data("world2"),
      aes(x = long,
          y = lat,
          group=group),
      fill = "grey50",
      color = "black") +
    coord_map() +
    geom_point(aes(x = x, y = y, 
                   fill = Index.categories.2020, 
                   size = Inclusiveness.index.2020),
               shape = 21,
               alpha = .75) +
    scale_fill_brewer(palette = "YlGnBu", direction = -1)
  
 

```

## Movies

```{r}

library(ggplot2movies) # add to index if you use this instead

# Movies were selected for inclusion if they had a known length and had 
# been rated by at least one imdb user.
movies <- movies

```

```{r}

movies %>% 
  pivot_longer(cols = c(Action, Animation, Comedy, Drama, Documentary, Romance, Short), names_to = "genre") %>%
  mutate(value = as.logical(value)) %>%
  dplyr::filter(value) %>%
  mutate(genre = genre %>%
           as_factor() %>%
           fct_infreq) %>%
  ggplot(aes(x=genre)) +
    geom_bar()
```

```{r}

ggplot(movies, aes(budget, rating)) +
  geom_point()

```

```{r}

movies %>%
  group_by(year) %>%
  summarise(mean_budget = mean(budget, na.rm=T)) %>%
  ggplot(aes(x=year, y=mean_budget)) +
    geom_line()

```

```{r}

movies %>% 
  pivot_longer(cols = c(Action, Animation, Comedy, Drama, Documentary, Romance, Short), names_to = "genre") %>%
  mutate(value = as.logical(value)) %>%
  dplyr::filter(value) %>%
  mutate(genre = genre %>%
           as_factor() %>%
           fct_infreq) %>%
  ggplot(aes(y="", fill=genre)) +
    geom_bar(position = position_fill()) +
    coord_polar(direction = -1)

```

```{r}

movies %>% 
  pivot_longer(cols = c(Action, Animation, Comedy, Drama, Documentary, Romance, Short), names_to = "genre") %>%
  mutate(value = as.logical(value)) %>%
  dplyr::filter(value, mpaa != "") %>%
  mutate(genre = genre %>%
           as_factor() %>%
           fct_infreq) %>%
  group_by(mpaa, genre) %>%
  summarize(n = n(), mean_budget = mean(budget, na.rm=T)) %>%
  ggplot(aes(x=mpaa, y=genre, fill = mean_budget)) +
    geom_tile()

```

```{r}

ggplot(movies, aes(x=rating)) +
  geom_histogram()

```

```{r}

movies %>% 
  pivot_longer(cols = c(Action, Animation, Comedy, Drama, Documentary, Romance, Short), names_to = "genre") %>%
  mutate(value = as.logical(value)) %>%
  dplyr::filter(value) %>%
  mutate(genre = genre %>%
           as_factor() %>%
           fct_infreq) %>%
  ggplot(aes(x=genre, y=rating)) +
    geom_boxplot()
  

```

(No maps, but maybe that's okay)
