```{r ch2-setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Building basic visualizations with ggplot2 {#building-basic-visualizations}

```{r, include=FALSE}

source('data/vineyards/load_vineyards.R')

```

In Chapter \@ref(reading-visualizations), we covered a variety of visualization types 
and how to read them. All
of those visualizations were created using an R package called `ggplot2`. This
package is a very common way of creating visualizations in R, partly because it is
so versatile. Once you understand the basics of how to create `ggplot2` visualizations, 
you can use the same techniques to produce an enormous range of visualizations.

In this chapter, we'll focus on the basic components of the R code you need to write
to create `ggplot2` visualizations. We'll use the word *syntax* to refer to the 
components you need to create and the way you need to organize those components.
Instead of trying to create a bunch of different visualizations, we start with one
simple visualization and keep adding to it to learn more about `ggplot2` features.
In future chapters we will cover other kinds of visualizations and more advanced
`ggplot2` features.

```{r, eval=FALSE}

ggplot(vineyards, aes(x=Visitors, y=CasePrice, color=Location, size = Acres)) +
  geom_point()

```


## Basic ggplot2 syntax

To learn the syntax of a `ggplot2` visualization, it helps to understand how
`ggplot2` thinks of a visualization (Figure \@ref(fig:ggplot-diagram)). 

```{r ggplot-diagram, echo=FALSE, fig.cap='The ggplot2 model of a visualization.'}

#TODO: move plot to the left and data to the right

ggplot(tibble()) + 
  geom_point() + 
  xlim(0, 100) + 
  ylim(38, 78) +
  coord_equal() +
  annotate("rect", xmin = 70, xmax = 100, ymin = 38, 
           ymax = 71, fill = "white", color = "black") +
  annotate("text", x = 85, y = 75, label = "plot", size = 8) +
  annotate("text", x = 85, y = 68, label = "geom", size = 6) +
  annotate("text", x = 85, y = 64, label = "(chart type or shape)", size = 3) +
  annotate("segment", x = 85, xend = 85, y = 62, yend = 52, 
           arrow = arrow(type = "closed", length = unit(.13, "inches"))) +
  annotate("text", x = 85, y = 49, label = "aesthetics", size = 6) +
  annotate("text", x = 85, y = 43, label = "(visual properties,\nlike x position, color)", 
           size = 3) +
  annotate("text", x = 54, y = 49, label = "mapping", size = 6) +
  annotate("segment", x = 63, xend = 74, y = 48.5, yend = 48.5, 
           arrow = arrow(type = "closed", length = unit(.13, "inches"))) +
  annotate("text", x = 17, y = 75, label = "data", size = 8) +
  annotate("rect", xmin = 0, xmax = 33, ymin = 38, 
           ymax = 71, fill = "white") +
  annotate("rect", xmin = 1, xmax = 8, ymin = 63, 
           ymax = 70, fill = "grey60") +
  annotate("rect", xmin = 9, xmax = 16, ymin = 63, 
           ymax = 70, fill = "grey60") +
  annotate("rect", xmin = 17, xmax = 24, ymin = 63, 
           ymax = 70, fill = "grey60") +
  annotate("rect", xmin = 25, xmax = 32, ymin = 63, 
           ymax = 70, fill = "grey60") +
  annotate("rect", xmin = 1, xmax = 8, ymin = 55, 
           ymax = 62, fill = "grey90") +
  annotate("rect", xmin = 9, xmax = 16, ymin = 55, 
           ymax = 62, fill = "grey90") +
  annotate("rect", xmin = 17, xmax = 24, ymin = 55, 
           ymax = 62, fill = "grey90") +
  annotate("rect", xmin = 25, xmax = 32, ymin = 55, 
           ymax = 62, fill = "grey90") +
  annotate("rect", xmin = 1, xmax = 8, ymin = 47, 
           ymax = 54, fill = "grey80") +
  annotate("rect", xmin = 9, xmax = 16, ymin = 47, 
           ymax = 54, fill = "grey80") +
  annotate("rect", xmin = 17, xmax = 24, ymin = 47, 
           ymax = 54, fill = "grey80") +
  annotate("rect", xmin = 25, xmax = 32, ymin = 47, 
           ymax = 54, fill = "grey80") +
  annotate("rect", xmin = 1, xmax = 8, ymin = 39, 
           ymax = 46, fill = "grey90") +
  annotate("rect", xmin = 9, xmax = 16, ymin = 39, 
           ymax = 46, fill = "grey90") +
  annotate("rect", xmin = 17, xmax = 24, ymin = 39, 
           ymax = 46, fill = "grey90") +
  annotate("rect", xmin = 25, xmax = 32, ymin = 39, 
           ymax = 46, fill = "grey90") +
  annotate("segment", x = 45, xend = 34, y = 48.5, yend = 48.5, 
           arrow = arrow(type = "closed", length = unit(.13, "inches"))) +
  theme_void()

```

In `ggplot2`, 
a visualization is called a "plot." Each plot uses one or more visualization types 
like the ones we learned in Chapter \@ref(reading-visualizations) (e.g., bar 
chart, scatter plot). The visualization type is called a "geom," which is short
for geometric object. You can think of the geom as the shape of the visual elements
in the plot.

Each possible geom has default visual properties. This makes sense if we step back
and think about a geom as a type of visualization. Every bar chart, for example, 
has a few things in common with every other bar chart: the numbers are represented 
by rectangles that vary in length, and there is a separate bar for each value of 
a categorical variable. To say that another way, a bar chart by default has one 
or more bars lined up along one axis, and the other axis allows you to compare 
the lengths of those bars. The default visual properties for a bar chart, then, 
are that it has two axes: one that serves to organize the bars, and one that 
serves to measure the lengths.

In `ggplot2`, the visual properties associated with each geom are called "aesthetics."
The aesthetics vary a bit from geom to geom, but there are some aesthetics that
are required for almost all geoms. The most common aesthetics are called "x" and "y",
and they stand for the x and y axes. Most geoms use an x axis to organize
one variable and a y axis to organize another. There are plenty of other optional 
aesthetics, however, that describe how a geom should look. The bars
in a bar chart don't just have a position on the x axis and a length on the y axis.
They also have a fill color, a fill transparency value, a border color, a border 
size, and a border line type. 

I think of a geom's aesthetics as questions on a form that I need to fill out for the
plot to work correctly. If something is option, like the fill color, I don't really
have to answer that question on the form. There is a default value `ggplot2` will
use if I don't have a specific value in mind. For the x and y axes, though, `ggplot2`
does not have a default. Those are required, and `ggplot2` will give me an error
if I try to make a plot without providing that information.

How do I tell `ggplot2` what values to use for the aesthetics, like the x and y 
axes? I need to create a "mapping" between the dataset I want to visualize and the
geom's aesthetics. Again, I think if we step back, this is pretty logical. We need
to get our data into our visualization. Different parts of the data need to be
associated with different parts of our visualization. That process of creating
associations between our dataset and our visualization is called mapping. In `ggplot2`,
we will map variables from our dataset to aesthetics in our geom one by one. For
each aesthetic we want to specify, we will pick a variable and map it to that
aesthetic. In a bar chart with vertical bars, we would map a categorical variable 
to the x aesthetic and a numerical variable to the y aesthetic. For horizontal bars,
the categorical variable would be mapped to the y aesthetic and the numerical
variable to the x aesthetic.

Looking again at our diagram in figure \@ref(fig:ggplot-diagram), we can summarize
it like this. A plot contains one or more geoms that specify the visualization type.
A geom determines the required and optional aesthetics for that part of the plot.
To bring data into the visualization, create a mapping between variables in a dataset
and the (required or options) aesthetics of the geom.

Now that we understand how `ggplot2` thinks about the structure of a plot, we can
turn to the code that we need to write.

### Simple Plot Template

In R, there is often many different ways of writing code to accomplish your goals.
The same is true in `ggplot2`. We'll start with learning a template for a simple
`ggplot2` plot, and then we will learn some additional ways of writing the code
that may be easier or work better for different kinds of visualizations.

```{r, echo=TRUE, eval=FALSE}

# start with the main plot function and set the data frame
ggplot(data = data_frame) + 
  
# next add a geometry layer and create the mapping between data frame and aesthetics
  geom_...(mapping = aes(aesthetics1 = data_field1, aesthetics2 = data_field))

```

## Example: Vineyards in the Finger Lakes Region of New York State

### Step 1: Add Main Plot Function

```{r ggpot-main-function, echo = TRUE}

ggplot()

```

### Step 2: Set the Data

```{r ggplot-set-data, echo = TRUE}

ggplot(data = vineyards)

```

### Step 3: Choose a shape layer

```{r ggplot-choose-shape, echo = TRUE, eval=FALSE}

# This will throw an error.
ggplot(data = vineyards) +
  geom_point()

# Error: geom_point requires the following missing aesthetics: x and y Run `rlang::last_error()` to see where the error occurred.

```

### Step 4: Map Variables to Aesthetics

```{r ggplot-map-aesthetics, echo = TRUE}

# Start with the required aesthetics: x and y.
# We'll be mapping x to the Visitors column and y to the CasePrice column.

mapped_variables <- aes(x=Visitors, y=CasePrice)

ggplot(data = vineyards) +
  geom_point(mapping = mapped_variables)

# or, all in one step:

ggplot(data = vineyards) +
  geom_point(mapping = aes(x=Visitors, y=CasePrice))


```

Now, we'll add a third variable. We'll also map the Location variable to the `color` aesthetic.

```{r ggplot-add-color, echo = TRUE}

# Start with the required aesthetics: x and y.
# We'll be mapping x to the Visitors column and y to the CasePrice column.

ggplot(data = vineyards) +
  geom_point(mapping = aes(x=Visitors, y=CasePrice, color=Location))


```

### Step 5: Adjust the Defaults

```{r ggplot-make-adjustments, echo = TRUE}

# Changing the default size of all of the points
ggplot(data = vineyards, mapping = aes(x=Visitors, y=CasePrice)) +
  geom_point(size=5)

# Changing the default color of all of the points
ggplot(data = vineyards, mapping = aes(x=Visitors, y=CasePrice)) +
  geom_point(color="blue")

# Changing both size and color
ggplot(data = vineyards, mapping = aes(x=Visitors, y=CasePrice)) +
  geom_point(size=5, color="blue")


```


## Inheritance

```{r}

ggplot(data = vineyards) +
  geom_point(mapping = aes(x=Visitors, y=CasePrice, color=Location)) +
  geom_text(mapping = aes(x=Visitors, y=CasePrice, color=Location, label=Vineyard),
            nudge_y=3)

ggplot(data = vineyards, mapping = aes(x=Visitors, y=CasePrice, color=Location)) +
  geom_point() +
  geom_text(mapping = aes(label=Vineyard), nudge_y=3, show.legend = F)

```

## How to Debug Code


