\cleardoublepage 

# (APPENDIX) Appendix {-}

# Datasets

```{r, include=FALSE}

source('data/load_all_data.R')

```

## Duke Enrollment {-}

[Duke enrollment](https://doi.org/10.7924/r4db82p1j)

Sample of Duke Enrollment By School dataset, Table \@ref(tab:duke-school-tab).

```{r duke-school-tab, echo=FALSE}

knitr::kable(
  head(duke_enrollment_by_school), caption = 'A sample from the Duke Enrollment By School dataset.',
  booktabs = TRUE
)

```


## Bar Chart

Figure \@ref(fig:duke-school-bar).

```{r duke-school-bar, fig.cap='Total Duke Enrollment by School', out.width='90%', echo=FALSE}

ggplot(duke_enrollment_by_school %>% 
         mutate(School = 
                  School %>% as_factor() %>% fct_infreq() %>% fct_rev()), 
       aes(y=School)) +
  geom_bar()

```


## Coral Resilience Data {-}

[Protecting coral reefs](https://doi.org/10.7924/G8348HFP)

Figure \@ref(fig:coral-resilience-point).

```{r coral-resilience-point, fig.cap='Log of tissue loss by snail density', out.width='90%', echo=FALSE}

# chart from p. 68 of https://hdl.handle.net/10161/16819 
#   (also published in https://doi.org/10.1038/s41559-018-0589-0)
ggplot(coral_resilience, aes(x=snails_exp, y=log(tissue_loss), color=treatment)) + geom_point()
```

```{r, include=FALSE}
# Could make a map with reef coordinates from the thesis (p.46):

#Site Name,Protection Status,Florida Keys Region,GPS Location
#Molasses Reef,SPA/ No-take,Upper Keys,N 25 0.579, W 80 22.471
#Conch Reef,SPA/ No-take,Upper Keys,N 24 57.110, W 80 27.564
#Coffins Patch,SPA/ No-take,Middle Keys,N 24 41.400, W 80 57.850
#Pickles Reef,Fishing allowed,Upper Keys,N 24 59.170, W 80 24.940
#Horseshoe Reef,Fishing allowed,Middle Keys,N 24 39.670, W 80 59.650
#American Shoals,Fishing allowed,Lower Keys,N 24 31.568, W 81 31.383

# but all of the experiment data comes from multiple colonies in one reef.
# p. 46 of thesis:

# "This experiment was conducted at Pickles Reef in Key Largo, Florida,
# USA from mid-June to mid-August 2014. Within the experimental site, I selected 
# 30 similarly-sized colonies (~159.1 ± 15.5 cm 2 surface area) that were already 
# harboring C. abbreviata."

```

## Git Experience {-}

[A Behavioral Approach to Understanding the Git Experience](https://osf.io/57tb8/)

```{r}

# won't work without aggregation
#ggplot(git_experience, aes(x=year_vcs)) +
#  geom_line()

ggplot(git_experience %>% count(year_vcs), aes(x=year_vcs, y=n)) +
  geom_line()

ggplot(git_experience %>% count(year_vcs, first_vcs), aes(x=year_vcs, y=n, color=first_vcs)) +
  geom_line()

```


```{r}

# First use vs. now use

first_used <- git_experience %>% 
  count(first_vcs) %>%
  rename("n_first_used" = "n", "system" = "first_vcs")

now_use <- git_experience %>%
  pivot_longer(cols=c(first_bazaar,first_cvs,first_git,
                      first_hg,first_monotone,first_svn, first_other),
               names_to = "system") %>%
  dplyr::filter(value == TRUE) %>%
  count(system) %>%
  rename("n_now_use" = "n") %>%
  mutate(system = case_when(
    system == "first_bazaar" ~ "Bazaar",
    system == "first_cvs" ~ "CVS",
    system ==  "first_git" ~ "Git",
    system ==  "first_hg" ~ "Mercurial",
    system ==  "first_monotone" ~ "Monotone",
    system ==  "first_svn" ~ "SVN", 
    system ==  "first_other" ~ "Other"
  ))

combined <- first_used %>% left_join(now_use) %>%
  pivot_longer(cols = c(n_first_used, n_now_use), 
               names_to = "time",
               values_to = "count") %>%
  drop_na()

ggplot(combined, aes(x=as_factor(time), y=count, group=system, color=system)) +
  geom_line()

```

```{r}

ggplot(git_experience, aes(x=first_vcs)) +
  geom_bar()

ggplot(git_experience %>%
         drop_na(first_vcs) %>%
         mutate(first_vcs = as_factor(first_vcs) %>% 
             fct_infreq()), 
         aes(x=first_vcs)) +
  geom_bar()


#How difficult was it for you to learn how to use git on your local computer?
# 1 - Very easy (1)
# 2 - Easy (2)
# 3 - Neutral (Neither easy nor difficult) (3)
# 4 - Difficult (4)
# 5 - Very difficult (5)

ggplot(git_experience %>% 
         drop_na(local_ease) %>%
         mutate(local_ease = as_factor(local_ease) %>% 
             fct_recode("Very easy"="1", 
                        "Easy"="2", "Neutral"="3", 
                        "Difficult"="4", "Very difficult"="5")), 
       aes(local_ease)) +
  geom_bar()

#How difficult was it for you to learn how to use the git hosting platform (e.g. GitLab, GitHub, etc.)?
# 1 - Very easy (1)
# 2 - Easy (2)
# 3 - Neutral (Neither easy nor difficult) (3)
# 4 - Difficult (4)
# 5 - Very difficult (5)
ggplot(git_experience %>% 
         drop_na(ghp_ease) %>%
         mutate(ghp_ease = as_factor(ghp_ease) %>% 
             fct_recode("Very easy"="1", 
                        "Easy"="2", "Neutral"="3", 
                        "Difficult"="4", "Very difficult"="5")), 
       aes(ghp_ease)) +
  geom_bar()

# How proficient do you think you are with git?
# 1 - Fundamental Awareness (basic knowledge) (1)
# 2 - Novice (limited experience) (2)
# 3 - Intermediate (practical application) (3)
# 4 - Advanced (applied theory) (4)
# 5 - Expert (recognized authority) (5)
ggplot(git_experience %>% 
         drop_na(proficiency) %>%
         mutate(proficiency = as_factor(proficiency) %>% 
             fct_recode("Basic Knowledge"="1", 
                        "Novice"="2", "Intermediate"="3", 
                        "Advanced"="4", "Expert"="5")), 
       aes(proficiency)) +
  geom_bar()

# How frequently do you have to reteach yourself git?
# Daily (1)
# Weekly (2)
# Once a semester (3)
# Once a year (4)
# Never (5)
# Other: (6)
ggplot(git_experience %>%
         drop_na(freq_reteach) %>%
         mutate(freq_reteach = as_factor(freq_reteach) %>% 
             fct_relevel("Daily", "Weekly", "Once a semester",
                         "Once a year", "Never", "Other:")), 
       aes(freq_reteach)) +
  geom_bar()



```

```{r}

git_experience %>% drop_na(first_vcs) %>%
  group_by(first_vcs) %>% 
  summarise(count = n(), mean_prof = mean(proficiency, na.rm=TRUE)) %>%
  ggplot(aes(x=first_vcs, y=mean_prof)) +
    geom_col()

git_experience %>% 
  drop_na(first_vcs) %>%
  group_by(first_vcs) %>% 
  summarise(count = n(), mean_proficiency = mean(proficiency, na.rm=TRUE)) %>%
  mutate(first_vcs = as_factor(first_vcs) %>% 
           fct_reorder(mean_proficiency) %>% fct_rev()) %>%
  ggplot(aes(x=first_vcs, y=mean_proficiency)) +
    geom_col()


```

```{r}

ggplot(git_experience, aes(x=year_vcs, y=first_vcs)) +
  geom_count()

git_experience %>%
  drop_na(year_vcs, first_vcs) %>%
  group_by(year_vcs, first_vcs) %>%
  summarize(min_year = min(year_vcs),
            count=n()) %>%
  mutate(first_vcs = as_factor(first_vcs) %>% 
           fct_reorder(min_year)) %>%
  ggplot(aes(x=year_vcs, y=first_vcs, size=count)) +
  geom_point()

git_experience %>%
  drop_na(first_vcs) %>%
  mutate(first_vcs = as_factor(first_vcs) %>% 
           fct_reorder(year_vcs, .fun=min)) %>%
  ggplot(aes(x=year_vcs, y=first_vcs)) +
  geom_count()

ggplot(git_experience %>% count(year_vcs, first_vcs), 
       aes(x=year_vcs, y=first_vcs, size=n)) +
  geom_point()

ggplot(git_experience, aes(x=year_vcs, y=first_vcs)) +
  geom_point()

ggplot(git_experience, aes(x=year_vcs, y=first_vcs, color=first_vcs)) +
  geom_jitter()

ggplot(git_experience %>% count(year_vcs, first_vcs), 
       aes(x=year_vcs, y=first_vcs, fill=n)) +
  geom_tile()

```

```{r}

ggplot(git_experience %>% drop_na(use_gitlab), aes(y="",fill=use_gitlab)) +
  geom_bar(position=position_fill()) +
  coord_polar()

```


## Inclusiveness Index {-}

[Inclusiveness Index](https://belonging.berkeley.edu/inclusivenessindex/data-and-resources)

```{r}

ggplot(inclusiveness_index, 
       aes(x = Z.score.General.Population, 
           y = Z.score.Gender,
           color = Continent)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE) + 
  facet_wrap(vars(Continent))

ggplot(inclusiveness_index, 
       aes(x = Z.score.General.Population, 
           y = Z.score.Sexual.Orientation)) +
  geom_point(aes(color = Continent)) +
  geom_smooth(method="lm", se=FALSE)

ggplot(inclusiveness_index, 
       aes(x = Z.score.General.Population, 
           y = Inclusiveness.index.2020)) +
  geom_point(aes(color = Continent)) +
  geom_smooth(method="lm", se=FALSE)

ggplot(inclusiveness_index, 
       aes(x = Z.score.Gender, 
           y = Z.score.Sexual.Orientation)) +
  geom_point(aes(color = Continent)) +
  geom_smooth(method="lm", se=FALSE)

ggplot(inclusiveness_index, 
       aes(x = Z.score.Gender, 
           y = Z.score.Religion)) +
  geom_point(aes(color = Continent)) +
  geom_smooth(method="lm", se=FALSE)

ggplot(inclusiveness_index, 
       aes(x = Z.score.Gender, 
           y = Inclusiveness.index.2020)) +
  geom_point(aes(color = Continent)) +
  geom_smooth(method="lm", se=FALSE)

ggplot(inclusiveness_index, 
       aes(x = Z.score.Sexual.Orientation, 
           y = Inclusiveness.index.2020)) +
  geom_point(aes(color = Continent)) +
  geom_text(data = inclusiveness_index %>% 
              dplyr::filter(Inclusiveness.index.2020 == max(Inclusiveness.index.2020, na.rm=TRUE) | Inclusiveness.index.2020 == min(Inclusiveness.index.2020, na.rm=TRUE)), aes(label = Country)) +
  geom_smooth(method="lm", se=FALSE)

### USE THIS ONE? ###
ggplot(inclusiveness_index, 
       aes(x = Z.score.Religion, 
           y = Inclusiveness.index.2020)) +
  geom_point(aes(color = Continent)) +
  geom_text(data = inclusiveness_index %>% 
              dplyr::filter(Inclusiveness.index.2020 == max(Inclusiveness.index.2020, na.rm=TRUE) | Inclusiveness.index.2020 == min(Inclusiveness.index.2020, na.rm=TRUE)), aes(label = Country)) + 
  geom_smooth(method="lm", se=FALSE)

ggplot(inclusiveness_index, 
       aes(group = Z.score.Disability, 
           y = Inclusiveness.index.2020)) +
  geom_boxplot()


```

## Candidate Demographics {-}

[Candidate Demographics](https://wholeads.us/research/rising-tide-ballot-demographics/)

Includes State, Candidate Name, Candidate Party, Office Name, White/Non-White, Race, Gender, Race/Gender Category, Office Level; 4 years (2012, 2014, 2016, 2018), over 40k records

## Affinity Spending {-}

[Affinity Spending](https://github.com/OpportunityInsights/EconomicTracker)

```{r}

ggplot(affinity_spending_cat, aes(x=date, y=spending_prop_vs_jan_2020, color=income_quartile)) +
  geom_line() +
  facet_wrap(vars(spending_category))

```

```{r}

affinity_spending_by_quartile <- affinity_spending_cat %>%
  pivot_wider(names_from = income_quartile, values_from = spending_prop_vs_jan_2020)

ggplot(affinity_spending_by_quartile %>% 
         dplyr::filter(spending_category == "retail_w_grocery"), 
       aes(x = q1, y = q4)) +
  geom_point()

```


```{r, fig.height=15}

affinity_spending_by_category <- affinity_spending_cat %>%
  pivot_wider(names_from = spending_category, values_from = spending_prop_vs_jan_2020)

ggplot(affinity_spending_by_category, 
       aes(x = durables, y = nondurables, color = income_quartile)) +
  geom_point() +
  facet_grid(rows = vars(paste0(year(date),"-", str_pad(month(date), 2, pad="0"))),
             cols = vars(income_quartile))
```


