\cleardoublepage 

# (APPENDIX) Appendix {-}

# Datasets

```{r, include=FALSE}

source('data/load_all_data.R')

```

## Duke Enrollment {-}

[Duke enrollment](https://doi.org/10.7924/r4db82p1j)

Sample of Duke Enrollment By School dataset, Table \@ref(tab:duke-school-tab).

```{r duke-school-tab, echo=FALSE}

knitr::kable(
  head(duke_enrollment_by_school), caption = 'A sample from the Duke Enrollment By School dataset.',
  booktabs = TRUE
)

```



## Coral Resilience Data {-}

[Protecting coral reefs](https://doi.org/10.7924/G8348HFP)

Figure \@ref(fig:coral-resilience-point).

```{r coral-resilience-point, fig.cap='Log of tissue loss by snail density', out.width='90%', echo=FALSE}

# chart from p. 68 of https://hdl.handle.net/10161/16819 
#   (also published in https://doi.org/10.1038/s41559-018-0589-0)
ggplot(coral_resilience, aes(x=snails_exp, y=log(tissue_loss), color=treatment)) + geom_point()
```

```{r, include=FALSE}
# Could make a map with reef coordinates from the thesis (p.46):

#Site Name,Protection Status,Florida Keys Region,GPS Location
#Molasses Reef,SPA/ No-take,Upper Keys,N 25 0.579, W 80 22.471
#Conch Reef,SPA/ No-take,Upper Keys,N 24 57.110, W 80 27.564
#Coffins Patch,SPA/ No-take,Middle Keys,N 24 41.400, W 80 57.850
#Pickles Reef,Fishing allowed,Upper Keys,N 24 59.170, W 80 24.940
#Horseshoe Reef,Fishing allowed,Middle Keys,N 24 39.670, W 80 59.650
#American Shoals,Fishing allowed,Lower Keys,N 24 31.568, W 81 31.383

# but all of the experiment data comes from multiple colonies in one reef.
# p. 46 of thesis:

# "This experiment was conducted at Pickles Reef in Key Largo, Florida,
# USA from mid-June to mid-August 2014. Within the experimental site, I selected 
# 30 similarly-sized colonies (~159.1 ± 15.5 cm 2 surface area) that were already 
# harboring C. abbreviata."

```

## Git Experience {-}

[A Behavioral Approach to Understanding the Git Experience](https://osf.io/57tb8/)


## Inclusiveness Index {-}

[Inclusiveness Index](https://belonging.berkeley.edu/inclusivenessindex/data-and-resources)

## Candidate Demographics {-}

[Candidate Demographics](https://wholeads.us/research/rising-tide-ballot-demographics/)

Includes State, Candidate Name, Candidate Party, Office Name, White/Non-White, Race, Gender, Race/Gender Category, Office Level; 4 years (2012, 2014, 2016, 2018), over 40k records

## Affinity Spending {-}

[Affinity Spending](https://github.com/OpportunityInsights/EconomicTracker)


