\cleardoublepage 

# (APPENDIX) Appendix {-}

# Appendix Placeholder

Placeholder text for the appendix.

To know more about **bookdown**, see https://bookdown.org.

## Possible datasets

**check again for licenses**

**need data good for text, repeated reports, interactive charts...** 

* [Duke enrollment](https://doi.org/10.7924/r4db82p1j)
* [Songbird color perception](https://doi.org/10.7924/r4rx96r99)
* [Wetlands and shorebirds](https://doi.org/10.7924/r47w6b882)
* [Better Baboon Breakups](https://doi.org/10.7924/r4bz67v25)
* [Electricity markets](https://doi.org/10.7924/r40k2dx36)
* [Truncating bar charts](https://osf.io/ytq3h/) - and related [paper](https://doi.org/10.1016/j.jarmac.2020.10.002)
* [Primate maternal death](https://doi.org/10.7924/r44f1tk8k)
* [Protecting coral reefs](https://doi.org/10.7924/G8348HFP)
* [Mountain rainfall](https://doi.org/10.7924/G8CJ8BJK)
* [Antitrust and preferential trade agreements](https://doi.org/10.7924/r4h70c24h)
* [Research vessel logs](https://doi.org/10.7924/r41j9fq8z)
* [SELF-PERCEPTIONS OF CREATIVITY & ARTS PARTICIPATION](https://doi.org/10.3886/ICPSR37853.v1)
* [Getty Provenance Index](https://github.com/thegetty/provenance-index-csv)
* [Investigating and Archiving the Scholarly Git Experience](https://doi.org/10.5064/F6VOIB8H) - CC-BY-SA-4.0
    * have to register for data, but maybe can use [GitLab dataset?](https://gitlab.com/investigating-archiving-git/survey-analysis/-/tree/main)
    * see also https://investigating-archiving-git.gitlab.io/updates/iasge-data-published/
* Wilkins, Luke (2021). Exploring the Emotional Well-being and Mental Health of Individuals who Play Fantasy Football, 2021. [Data Collection]. Colchester, Essex: UK Data Service. [10.5255/UKDA-SN-854823](https://dx.doi.org/10.5255/UKDA-SN-854823)
* Meadows, Angela (2020). Weight stigma resistance scale validation study 2018-2019. [Data Collection]. Colchester, Essex: UK Data Service. [10.5255/UKDA-SN-854091](https://dx.doi.org/10.5255/UKDA-SN-854091)
* Meadows, Angela (2020). A social identity model of weight stigma resistance in high-weight individuals 2015. [Data Collection]. Colchester, Essex: UK Data Service. [10.5255/UKDA-SN-854100](https://dx.doi.org/10.5255/UKDA-SN-854100)
* Meadows, Angela (2020). Workplace wellbeing study 2018-2019. [Data Collection]. Colchester, Essex: UK Data Service. [10.5255/UKDA-SN-854102](https://dx.doi.org/10.5255/UKDA-SN-854102)
* Mulugeta, Mercy and Pertaub, David-Paul and Stevenson, Jed and Tekle Loyale, Dessalegn (2019). Retrospective household and livelihoods survey in Nyangatom District, South Omo, Ethiopia 2019. [Data Collection]. Colchester, Essex: UK Data Service. [10.5255/UKDA-SN-854024](https://dx.doi.org/10.5255/UKDA-SN-854024)
* http://nesstar.ukdataservice.ac.uk/webview/index.jsp?v=2&mode=documentation&submode=abstract&study=http://nesstar.ukdataservice.ac.uk:80/obj/fStudy/7913&top=yes
* http://nesstar.ukdataservice.ac.uk/webview/index.jsp?v=2&mode=documentation&submode=abstract&study=http://nesstar.ukdataservice.ac.uk:80/obj/fStudy/8150&top=yes
* https://www.icpsr.umich.edu/web/appfed/search/studies
* https://www.icpsr.umich.edu/web/RCMD/search/studies?start=0&sort=score%20desc%2CTITLE_SORT%20asc&RESTRICTED_DATA_TYPE=Public%20Use&ARCHIVE=RCMD&PUBLISH_STATUS=PUBLISHED&FORMAT_FACET=Delimited&rows=50&q=
* https://www.openicpsr.org/openicpsr/search/studies?start=0&sort=DATEUPDATED%20desc&GEOGRAPHY_FACET=United%20States&RESTRICTED_DATA_TYPE=Public%20Use&ARCHIVE=openicpsr&rows=346
* https://www.icpsr.umich.edu/web/DAIRL/search/studies?start=0&ARCHIVE=DAIRL&PUBLISH_STATUS=PUBLISHED&RESTRICTED_DATA_TYPE=Public%20Use&sort=TITLE_SORT%20asc&rows=50
* https://belonging.berkeley.edu/inclusivenessindex/data-and-resources
* https://www.nsf.gov/statistics/srvyrecentgrads/#tabs-2
* https://www.ers.usda.gov/data-products/food-access-research-atlas/download-the-data/
* http://millionsongdataset.com/pages/getting-dataset/
* https://www.sciencebase.gov/catalog/item/57bdfd8fe4b03fd6b7df5ff9
* https://w2.weather.gov/climate/
* https://vincentarelbundock.github.io/Rdatasets/datasets.html
* https://github.com/higgi13425/medicaldata
* https://wholeads.us/datasets/
    * https://wholeads.us/research/rising-tide-ballot-demographics/ - includes State, Candidate Name, Candidate Party, Office Name, White/Non-White, Race, Gender, Race/Gender Category, Office Level; 4 years (2012, 2014, 2016, 2018), over 40k records
* https://reports.nlihc.org/oor/alabama
* https://nhc.org/paycheck-to-paycheck/
* https://dsl.richmond.edu/panorama/redlining/#loc=5/39.1/-94.58&text=downloads
* https://tracktherecovery.org/
* https://www.epi.org/data/
* https://livingwage.mit.edu/
* https://www.epi.org/resources/budget/

** to consider: what kind of data would people see in different business
contexts? as opposed to research projects... **

* orders/product usage
* website usage
* satisfaction surveys
* donations in response to media campaigns
* email click rate? other engagement in relation to media campaigns?
* something for health? trying to predict/understand types of human behavior?
* 

## Duke Enrollment Data

```{r, include=FALSE}

library(tidyverse)

```

```{r}

duke_enrollment <- read_csv('data/duke_enrollment/UberMegaMaster_70S-20F-V5.csv') %>%
  dplyr::select(-X1, -X.2, -X.1, -X)

duke_enrollment_by_status <- duke_enrollment %>%
  dplyr::select(Year, Semester, Origin, Region, Sex, All_UG, All_Grad) %>%
  rename("Undergraduate" = "All_UG", "Graduate" = "All_Grad") %>%
  pivot_longer(cols = c(Undergraduate, Graduate), 
               names_to = "Student_Status", 
               values_to = "Count") %>%
  dplyr::filter(Count > 0)

duke_enrollment_by_school <- duke_enrollment %>%
  dplyr::select(Year, Semester, Origin, Region, Sex, Trinity, Nursing, 
                Engineering, Graduate, Divinity, Law, Business, Environment, 
                Medicine, Graduate.Nursing, Allied.Health) %>%
  pivot_longer(cols = c(Trinity, Nursing, Engineering, Graduate, Divinity, Law, 
                        Business, Environment, Medicine, Graduate.Nursing, 
                        Allied.Health), names_to = "School", values_to = "Count") %>%
  dplyr::filter(Count > 0)

write_csv(duke_enrollment_by_status, "data/duke_enrollment/duke_enrollment_by_status.csv")
write_csv(duke_enrollment_by_school, "data/duke_enrollment/duke_enrollment_by_school.csv")


```

